---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# STRINGDatabaseManipulation

The goal of STRINGDatabaseManipulation is to provide functions for reading in and manipulating [STRING](https://string-db.org/) protein-protein interaction networks.

## Installation

Only the development version of STRINGDatabaseManipulation is available. 

``` r
remotes::install_github("moseleyBioinformaticsLab/STRINGDatabaseManipulation")
```

## Issues

If you have a question about how to use the package, a request for something new to be implemented, or have a bug, please [file an issue](https://github.com/moseleyBioinformaticsLab/STRINGDatabaseManipulation/issues).

## Getting Data

This package works with preprocessed data from STRING itself. 

You can preprocess files from STRING by downloading link detail and alias files.
These files are available from the [STRING download page](https://string-db.org/cgi/download).
I **do not** recommend trying to work with the **full** interaction network across all species, but rather download files for your specific species of interest.
For example, you can filter to just the [human files](https://string-db.org/cgi/download?&species_text=Homo+sapiens).

We can get the files we want at the command line using a tool like `wget`:

```
# get the PPI network itself
wget https://stringdb-static.org/download/protein.links.detailed.v11.5/9606.protein.links.detailed.v11.5.txt.gz
# get aliases
wget https://stringdb-static.org/download/protein.aliases.v11.5/9606.protein.aliases.v11.5.txt.gz
``` 

And then process them so they are easier to use:

```{r process_data, eval = FALSE}
library(STRINGDatabaseManipulation)
# assuming data is in the file, you would do:
ppi_data = process_string_links("9606.protein.links.detailed.v11.5.txt.gz")

# save it for later
saveRDS(ppi_data, file = "tmp_ppi_data.rda")

# process aliases
protein_aliases = process_string_aliases("9606.protein.aliases.v11.5.txt.gz")

# save for later
saveRDS(protein_aliases, file = "tmp_aliases.rda")
```

This package has smaller versions of the link and data files (10,000 links) saved for examples, both as raw text files, and as package data files.
These files have been filtered to those links with an **experimental** score >= 400, as well as links that include EGFR and TP53.

```{r example_data}
library(STRINGDatabaseManipulation)
links_file = system.file("extdata", "STRING11.5_9606_links_raw.txt.gz", package = "STRINGDatabaseManipulation")
ppi_data = process_string_links(links_file)
head(ppi_data)

aliases_file = system.file("extdata", "STRING11.5_9606_aliases_raw.txt.gz", package = "STRINGDatabaseManipulation")
ppi_aliases = process_string_aliases(aliases_file)
head(ppi_aliases)
```

## Using Data

Now lets actually do something with the STRING data.
The most common analysis we want to do is find proteins that interact (directly or indirectly) with one or more query proteins.

We will use the example data from the package.

To filter the data, we can use `dplyr` to choose which evidence or set of evidences to use a filter.
Here we will use `combined_score` >= 400 (it's actually already filtered, but this is to show how we can use it).

```{r}
#| label: filter_ppi
ppi_filtered = ppi_data |>
  dplyr::filter(combined_score >= 400)
dim(ppi_data)
dim(ppi_filtered)
```

In this case it doesn't change the number of interactions, because the data was pre-filtered to make it tractable for inclusion in the package.

## TP53

Lets take everyone's favorite cancer gene, TP53, and look for those proteins that experimentally are known to interact with it.
We will find the STRING-db ID, and then query everything that is connected to it.

```{r}
#| label: find_tp53
ppi_graph = string_2_tidygraph(ppi_data)
tp53_alias = ppi_aliases |>
  dplyr::filter(other %in% "TP53")
tp53_alias
```

Now we can go fetch the neighbors of our query protein (n_hops = 0), and find everything that it interacts with.
Notice, `n_hops = 0`!
This is because the `hops` refers to hops over other proteins.
To find just the interacting pairs, we are doing **0** hops.

```{r}
#| label: tp53_interactions
tp53_interactions = find_nodes_n_hops(ppi_graph, n_hops = 0, start_nodes = tp53_alias$string)
tp53_interactions$graph
```

Here we can see that the returned interactions with TP53 includes another `r tp53_interactions$graph |> tidygraph::activate(nodes) |> tibble::as_tibble() |> nrow() - 1` proteins.
This is in **stark** contrast to the **maximum of 50** first shell entries returned by the STRING web tool.
In addition to the large number of interacting proteins, notice that many of them are connected to each other as well, given the number of edges compared to the number of nodes.


We can also examine, for this limited network, how many genes TP53 interacts with when we allow one protein in between.
For this example, this is likely to be lower than the real number, as it is a limited dataset.
In reality, TP53 interactors with a single hop is an incredibly large number of interactors that you will want to watch your memory usage.

```{r}
#| label: tp53_onehop
tp53_onehop = find_nodes_n_hops(ppi_graph, n_hops = 1, start_nodes = tp53_alias$string)
tp53_onehop$graph
```

Even with the limited network, we end up with a whopping `r tp53_onehop$graph |> tidygraph::activate(nodes) |> tibble::as_tibble() |> nrow() - 1` proteins that interact one hop out from TP53!

## To Do

There are several improvements we have in mind to make to make this package more useful:

* [functions to easily add protein identifiers to the returned network.](https://github.com/MoseleyBioinformaticsLab/STRINGDatabaseManipulation/issues/3)
* [take pathway data from the `graphite` package and work with it, we aren't actually limited to STRING data.](https://github.com/MoseleyBioinformaticsLab/STRINGDatabaseManipulation/issues/5)
* [visualization examples.](https://github.com/MoseleyBioinformaticsLab/STRINGDatabaseManipulation/issues/2)
* [tests. We really need more tests.](https://github.com/MoseleyBioinformaticsLab/STRINGDatabaseManipulation/issues/1)
* [could we use `sqlite` databases to enable analysis of things with **really** large numbers of interactors?](https://github.com/MoseleyBioinformaticsLab/STRINGDatabaseManipulation/issues/4)

## Code of Conduct

Note that the 'STRINGDatabaseManipulation' project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

